---
title: 'Regression Models Course Project'
author: "Gabriel Quintanar"
output:
  html_document:
    df_print: paged
---

```{r setoptions, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align='center', fig.width=10)
library(plyr); library(dplyr)
library(ggplot2); library(knitr)
```

## Executive Summary
This research will show the analysis on the dataset mtcars and address two objectives:
      - Is an automatic or manual transmission better for MPG?
      - Quantify the MPG difference between automatic and manual transmissions
      
In order to complete the task, some regression analysis will be made to accomplish the objectives.     
      
      
## Loading and Exploratory Analysis

Documentation about the dataset:

### Motor Trend Car Road Tests

Description

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

Usage

_mtcars_

Format

A data frame with 32 observations on 11 variables.

1.	 mpg	 Miles/(US) gallon
2.	 cyl	 Number of cylinders
3.	 disp	 Displacement (cu.in.)
4.	 hp	 Gross horsepower
5.	 drat	 Rear axle ratio
6.	 wt	 Weight (1000 lbs)
7.	 qsec	 1/4 mile time
8.	 vs	 V/S
9.	 am	 Transmission (0 = automatic, 1 = manual)
10.	 gear	 Number of forward gears
11.	 carb	 Number of carburetors


Here is a summary and the head of _mtcars_ data
```{r loading}
data("mtcars")
kable(head(mtcars, 5))
summ <- summary(mtcars)
kable(summ)
mtcars$am <- factor(mtcars$am)
levels(mtcars$am)[1] <- "Automatic"
levels(mtcars$am)[2] <- "Manual"
nManual <- mtcars %>% filter(am == "Manual") %>% count()
nAuto <- mtcars %>% filter(am == "Automatic") %>% count()
Transmission <- levels(mtcars$am)
tabl <- data.frame(Transmission, Counts = c(nAuto$n, nManual$n))
kable(tabl)
```

Our main objectives only needs variables _mpg_ and _am_. Checking the distribution those variables with graphics may lead to a better understanding of the data.


With this histograms showed in the figure 1, we see how cars with manual transmission has higher mpg records.

```{r}
meanAuto <- mtcars %>% filter(am == "Automatic") %>% summarise(mean = mean(mpg))
meanManual <- mtcars %>% filter(am == "Manual") %>% summarise(mean = mean(mpg))
kable(data.frame(Transmission, Means = c(meanAuto$mean, meanManual$mean)))
```

The mean, as shown in the table suggest a higher value for manual transmission. Figure 2 shows this exactly.


## Regression Analysis

Lets create a model with mpg as the outcome and am as the predictor.

```{r}
fit<- lm(mpg ~ factor(am), mtcars)
summary(fit)
confint(fit)
```

As seen in the summary. The increase in mpg is higher with a manual transmission. But considering the other variables, it may not be that way, as suggested by the Adjusted R-squared of 33.85%. Now lets compare the other variables.

```{r}
fit6 <- update(fit, mpg ~ am + cyl + disp + hp + drat + wt + qsec + vs + gear + carb)
fit5 <- update(fit, mpg ~ am + cyl + disp + hp + drat + wt + qsec + vs + gear)
fit4 <- update(fit, mpg ~ am + cyl + disp + hp + drat + wt + qsec + vs)
fit3 <- update(fit, mpg ~ am + cyl + disp + hp + drat + wt + qsec)
fit2 <- update(fit, mpg ~ am + cyl + disp + hp + wt)

anova(fit, fit2, fit3, fit4, fit5, fit6)

```

After comparing different variable in the model it is clear the the second model _(lm = mpg ~ am + cyl + disp + hp + wt)_ has the more significant p-value.

```{r}
summary(fit2)
```

Now, our Adjusted R-squared value is 82.73%. Now we can be sure that our model explains 85.51% of the variance. Plots for correlation between variables are shown in figure 3 and residual plots in figure 4.


## Conclusion


So, having the model _mpg ~ am + cyl + disp + hp + wt_, we can expect an increase of 1.55 MPG, holding the other variables constant when the transmission is manual. However, variables like cyl, hp and wt can decrease it, logically.

- Is an automatic or manual transmission better for MPG?
Acording our model, manual transmission is better for MPG.

- Quantify the MPG difference between automatic and manual transmissions
_1.55 MPG increase holding cyl, wt, disp and hp constant._


## Apendix

- Figure 1
```{r figure1}
mtcars %>% ggplot(aes(x = mpg, fill = am)) +
      geom_histogram(color = "black") +
      facet_grid(. ~ am) +
      labs(title = "Manual vs Automatic") +
      guides(fill = "none", color = "none")
```

- Figure 2
```{r figure2}
mtcars %>% ggplot(aes(x = am, y = mpg, color = am)) +
      geom_boxplot()
```

- Figure 3
```{r figure 3}
library(GGally)
modelData <- mtcars %>% dplyr::select(mpg, am, cyl, disp, hp, wt)
ggpairs(modelData, lower = list(continuous = "smooth"), wrap = c(method = "loess"))
```

- Figure 4
```{r}
par(mfrow = c(2,2))
plot(fit2)
```

