---
title: 'Statistical Inference Project Part 2: Basic Inferential Data Analysis'
author: "Gabriel Quinmat"
date: "6 de mayo de 2018"
output:
  html_document:
    df_print: paged
---

```{r setoptions, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.align='center', fig.width=10)
library(plyr); library(dplyr)
library(ggplot2);
```

## Overview
Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

Load the ToothGrowth data and perform some basic exploratory data analyses
Provide a basic summary of the data.
Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
State your conclusions and the assumptions needed for your conclusions.

## Loading and Analysis


We are going to work with ToothGrow dataset. The Effect of Vitamin C on Tooth Growth in Guinea Pigs.
```{r loadingData}
library("datasets")
data("ToothGrowth")
summary(ToothGrowth)
```


Looking the data we know now that we have 60 observations, in which 30 correspond to guinea pigs that were given vitamin c and the others Orange Juice. We can separate them in differents datasets to analyze them.

```{r separating}
ojT <- ToothGrowth %>% filter(supp == "OJ") %>% select(len, dose)
vcT <- ToothGrowth %>% filter(supp == "VC") %>% select(len, dose)
```

A quick summary of both datasets:


For supp OJ:
```{r} 
summary(ojT)
```


For supp VC:
```{r} 
summary(vcT)
```



```{r quickPlots}
ToothGrowth %>% ggplot(aes(x = dose, y = len, fill = supp)) +
      geom_bar(position = "dodge", stat = "identity") +
      labs(title = "Tooth Length ")
```

So we can see how tooth length is increasing as supplement dose increases. At first, vitamin C is not effecttive as orange juice, but with 2 mm/day is greater.

```{r statValues}
muOj <- mean(ojT$len)
muVc <- mean(vcT$len)

varOj <- var(ojT$len)
varVc <- var(vcT$len)

sdOj <- sd(ojT$len)
sdVc <- sd(vcT$len)
```

Orange Juice Length Mean: `r muOj`

Orange Juice Length Variance: `r varOj`

Orange Juice Length Standard Deviation: `r sdOj`



Vitamin C Length Mean: `r muVc`

Vitamin C Length Variance: `r varVc`

Vitamin C Length Standard Deviation: `r sdVc`



## Hypothesis Test & Confidence Intervals

Gosset's distribution will be a good distribution for testing as the dataset is not large. 

For making t test, we need to get the difference between both supplements. Let's make a hypothesis

H0 <- muOj - muVc = 0
Ha <- muOj - muVc != 0


### Length tooth test
```{r lengthTest}
test <- t.test(ojT$len, vcT$len, paired = TRUE)
```


Now he know the confidence interval:
`r test$conf.int`


Interval does not cross zero.

And a P-Value:
`r test$p.value`


And P-Value is close to 0.


Thus, we can see that there is a difference between supplement and we can reject the null hypothesis.


### Dose and Length test

We will select only 1.0 and 2.0 doses.
```{r dlTest}
toothDose <- ToothGrowth %>% filter(dose == 1.0 | dose == 2.0)
test2 <- t.test(len~dose, paired = TRUE, data = toothDose)
```

Confidence interval:
`r test2$conf.int`


And P-Value:
`r test2$p.value`


Again, but more significantly, we have a confidence interval that does not across zero and a p-value close to zero. so we can reject the null hypothesis.
